\chapter*{List of Abbreviations}


\textbf{AES} -- Advanced encryption standard.
\vspace{0.5cm}\\
\textbf{ALSZ13} -- OT extension protocol developed by Gilad Asharov, Yehuda Lindell, Thomas Schneider, and Michael Zohner \cite{ALSZ13}.
\vspace{0.5cm}\\
\textbf{API} -- Application programming interface.
\vspace{0.5cm}\\
\textbf{BB84} -- Quantum key distribution protocol developed by Charles Bennet and Gilles Brassard in 1984 \cite{BB84}.
\vspace{0.5cm}\\
\textbf{BBCS} -- Quantum oblivious transfer protocol developed by Bennet, Brassard, Crépeau and Skubiszewska \cite{BBCS92}.
\vspace{0.5cm}\\
\textbf{BCJL} -- Quantum bit commitment protocol developed by Brassard, Crépeau, Jozsa and Langlois \cite{BCJL93}.
\vspace{0.5cm}\\
\textbf{BGW} -- Secure multiparty computation protocol developed by Ben-Or, Goldwasser and Wigderson \cite{BGW88}.
\vspace{0.5cm}\\
\textbf{BM} -- Oblivious transfer protocol developed by Bellare and Micali \cite{BM89}.
\vspace{0.5cm}\\
\textbf{BMR} -- Secure multiparty computation protocol developed by Beaver, Micali and Rogaway \cite{BMR90}.
\vspace{0.5cm}\\
\textbf{BQS} -- Bounded-quantum-storage model.
\vspace{0.5cm}\\
\textbf{BQS-UC} -- Bounded-quantum-storage universal composability model.
\vspace{0.5cm}\\
\textbf{CBMC-GC} -- C Bounded model checker - Garbled Circuit.
\vspace{0.5cm}\\
\textbf{CCD} -- Secure multiparty computation protocol developed by Chaum, Crépeau and Damg\'{a}rd \cite{CCD88}.
\vspace{0.5cm}\\
\textbf{COM} -- Commitment.
\vspace{0.5cm}\\
\textbf{CP} -- Completely positive map.
\vspace{0.5cm}\\
\textbf{CPTP} -- Completely positive trace preserving map.
\vspace{0.5cm}\\
\textbf{CPU} -- Central processing unit.
\vspace{0.5cm}\\
\textbf{CRS} -- Common reference string.
\vspace{0.5cm}\\
\textbf{CSRNG} -- Cryptographically secure pseudorandom number generator.
\vspace{0.5cm}\\
\textbf{DI} -- Device independent.
\vspace{0.5cm}\\
\textbf{EGL} -- Even, Goldreich, Lempel.
\vspace{0.5cm}\\
\textbf{F84} -- Evolutionary distance developed by Felsenstein.
\vspace{0.5cm}\\
\textbf{FSA} -- Faked-state attacks.
\vspace{0.5cm}\\
\textbf{GDPR} -- General data protection regulation.
\vspace{0.5cm}\\
\textbf{GISAID} -- Global initiative on sharing avian influenza data.
\vspace{0.5cm}\\
\textbf{GMW} -- Secure multiparty computation protocol developed by Goldreich, Micali and Wigderson \cite{GMW87}.
\vspace{0.5cm}\\
\textbf{GWAS} -- Genome-wide association studies.
\vspace{0.5cm}\\
\textbf{HyCC} -- Compilation of hybrid protocols developed in \cite{Bscher2018}.
\vspace{0.5cm}\\
\textbf{JC} -- Jukes-Cantor.
\vspace{0.5cm}\\
\textbf{K2P} -- Kimura 2-parameter.
\vspace{0.5cm}\\
\textbf{KOS15} -- OT extension protocol developed by Keller, Orsini and Scholl \cite{KOS15}.
\vspace{0.5cm}\\
\textbf{LAN} -- Local area network.
\vspace{0.5cm}\\
\textbf{LD} -- LogDet.
\vspace{0.5cm}\\
\textbf{LWE} -- Learning with errors.
\vspace{0.5cm}\\
\textbf{M-LWE} -- Module learning with errors.
\vspace{0.5cm}\\
\textbf{MDI} -- Measurement device independent.
\vspace{0.5cm}\\
\textbf{MMH} -- Multi-linear modular hashing.
\vspace{0.5cm}\\
\textbf{MUB} -- Mutually unbiased bases.
\vspace{0.5cm}\\
\textbf{NP} -- Oblivious transfer protocol developed by Naor and Pinkas \cite{NP01}.
\vspace{0.5cm}\\
\textbf{NQS} -- Noisy quantum storage model.
\vspace{0.5cm}\\
\textbf{NTRU} -- Number theory research unit.
\vspace{0.5cm}\\
\textbf{OKM} -- Oblivious key management system.
\vspace{0.5cm}\\
\textbf{OLE} -- Oblivious linear evaluation.
\vspace{0.5cm}\\
\textbf{OT} -- Oblivious transfer.
\vspace{0.5cm}\\
\textbf{PDQ} -- Private database queries.
\vspace{0.5cm}\\
\textbf{PET} -- Privacy-enhancing technologies.
\vspace{0.5cm}\\
\textbf{PHYLIP} -- Phylogeny inference package.
\vspace{0.5cm}\\
\textbf{POVM} -- Positive operator-valued measure.
\vspace{0.5cm}\\
\textbf{PRG} -- Pseudorandom generator.
\vspace{0.5cm}\\
\textbf{QBC} -- Quantum bit commitment.
\vspace{0.5cm}\\
\textbf{QKD} -- Quantum key distribution.
\vspace{0.5cm}\\
\textbf{QOKD} -- Quantum oblivious key distribution.
\vspace{0.5cm}\\
\textbf{QOLE} -- Quantum oblivious linear evaluation.
\vspace{0.5cm}\\
\textbf{QOT} -- Quantum oblivious transfer.
\vspace{0.5cm}\\
\textbf{QRNG} -- Quantum random number generator.
\vspace{0.5cm}\\
\textbf{quantum-UC} -- Quantum universal composability model.
\vspace{0.5cm}\\
\textbf{RNG} -- Random number generator.
\vspace{0.5cm}\\
\textbf{RSA} -- Public-key cryptosystem developed by Rivest, Shamir and Adleman \cite{RSA78}.
\vspace{0.5cm}\\
\textbf{RWOLE} -- Random weak oblivious linear evaluation.
\vspace{0.5cm}\\
\textbf{SARS-CoV-2} -- Severe acute respiratory syndrome coronavirus 2.
\vspace{0.5cm}\\
\textbf{SHA} -- Secure hash algorithm.
\vspace{0.5cm}\\
\textbf{SMC} -- Secure multiparty computation.
\vspace{0.5cm}\\
\textbf{THA} -- Trojan-horse attack.
\vspace{0.5cm}\\
\textbf{UC} -- Universal composability model.
\vspace{0.5cm}\\
\textbf{UPGMA} -- Unweighted pair group method with arithmetic mean.
\vspace{0.5cm}\\
\textbf{VM} -- Virtual machine.
\vspace{0.5cm}\\
\textbf{VOLE} -- Vector oblivious linear evaluation.
\vspace{0.5cm}\\
\textbf{WOLE} -- Weak oblivious linear evaluation.
\vspace{0.5cm}\\
\textbf{WSE} -- Weak string erasure.
\vspace{0.5cm}\\








